### 2017-10-31 submission/manuscript.Rmd
add results section 2

@@

***Genera Previously Associated with Carcinoma Increases Relative Risk More than Alpha Diversity:***

>>>>

#Load in relative risk select genera data tables
select_RR_stool <- as.data.frame(read_csv("../data/process/tables/select_genus_RR_stool_composite.csv"))
select_RR_tissue <- as.data.frame(read_csv("../data/process/tables/select_genus_RR_tissue_composite.csv"))

@@

***Genera Previously Associated with Carcinoma Increases Relative Risk More than Alpha Diversity:***
Both fecal and tissue samples had a significantly increased RR for carcinoma but not for adenoma [Figure 3] which was greater than either evenness or Shannon diversity [Figure 2 & 3]. The relative risk did not increase when considering the total abundance or increasing number of carcinoma associated genera [Figure 3]. The RR effect size was greater for stool (RR range = `r paste(round(min(filter(select_RR_stool, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))` - `r paste(round(max(filter(select_RR_stool, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))`) then for tissue (RR range = `r paste(round(min(filter(select_RR_tissue, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))` - `r paste(round(max(filter(select_RR_tissue, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))`) . This decrease may be explained by the fact that tissue samples include matched samples.

### 2017-10-31 submission/manuscript.tex
add results section 2

\textbf{\emph{Fecal Diversity is Lower in Those with Carcinoma and
increases Relative Risk:}} Using power transformed and Z-score
normalized alpha diversity metrics both evenness and the Shannon
diversity metrics in feces are lower in those with carcinoma then in
controls but not for tissue samples {[}Figure 1{]}. Using linear
mixed-effects to control for study and variable region there was a
significant decrease from control to adenoma to carcinoma for both
evenness (P-value = 0.025) and Shannon diversity (P-value = 0.043). This
effect was not observed in tissue when additionally controlling for
whether the sample came from the same individual (P-value \textgreater{}
0.05). For fecal samples a decrease in Shannon diversity and evenness
resulted in a significantly increased relative risk for carcinoma
(P-value = 0.01 and P-value = 0.0011, respectively) {[}Figure 2{]}.
Although these values were significant the effect size was relatively
small for both metrics (Shannon RR = 1.31 and evenness RR = 1.34)
{[}Figure 2{]}. There was no increased relative risk for these metrics
for adenoma or for tissue in general {[}Figure S1-3{]}

\textbf{\emph{Section 2}}

>>>>

\textbf{\emph{Fecal Diversity is Lower in Those with Carcinoma and
Increases Relative Risk:}} Using power transformed and Z-score
normalized alpha diversity metrics both evenness and the Shannon
diversity metrics in feces are lower in those with carcinoma then in
controls but not for tissue samples {[}Figure 1{]}. Using linear
mixed-effects to control for study and variable region there was a
significant decrease from control to adenoma to carcinoma for both
evenness (P-value = 0.025) and Shannon diversity (P-value = 0.043). This
effect was not observed in tissue when additionally controlling for
whether the sample came from the same individual (P-value \textgreater{}
0.05). For fecal samples a decrease in Shannon diversity and evenness
resulted in a significantly increased relative risk for carcinoma
(P-value = 0.01 and P-value = 0.0011, respectively) {[}Figure 2{]}.
Although these values were significant the effect size was relatively
small for both metrics (Shannon RR = 1.31 and evenness RR = 1.34)
{[}Figure 2{]}. There was no increased relative risk for these metrics
for adenoma or for tissue in general {[}Figure S1-3{]}

\textbf{\emph{Genera Previously Associated with Carcinoma Increases
Relative Risk More than Alpha Diversity:}} Both fecal and tissue samples
had a significantly increased RR for carcinoma but not for adenoma
{[}Figure 3{]} which was greater than either evenness or Shannon
diversity {[}Figure 2 \& 3{]}. The relative risk did not increase when
considering the total abundance or increasing number of carcinoma
associated genera {[}Figure 3{]}. The RR effect size was greater for
stool (RR range = 1.78 - 2.64) then for tissue (RR range = 1.33 - 1.53)
. This decrease may be explained by the fact that tissue samples include
matched samples.
